### 2021-06-24 introduction.tex

Almost two-thirds of the world use mobile technologies~\cite{Comscore}, and the Android Operating System has dominated the market of smartphones, tablets, and others electronic devices \cite{statcounter}. Due to this growing popularity, the number of incidents related to Android malicious software (malware) has significantly increased~\cite{DBLP:journals/comsur/FarukiBLGGCR15,DBLP:journals/csur/SufatrioTCT15}\kn{Can we here say by how much?}. Security issues in Android software applications~\footnote{In this paper, we will use the terms Android Applications, Android Apps and Apps interchangeably, to represent Android software applications} have become a relevant research topic, and many techniques have been developed to identify vulnerabilities in Android apps~\cite{DBLP:conf/pldi/ArztRFBBKTOM14}, including the use of static analysis algorithms either to identify privacy leaks or to reveal the misuse of cryptographic primitives~\cite{krueger:ecoop-2018,rahaman:ccs-2019}, for instance.

Another perspective to the solving the same problem are approaches that rely on the use of dynamic analysis to mine Android sandboxes~\cite{DBLP:conf/icse/JamrozikSZ16}. With dynamic approaches, during an exploratory phase, one leverages an automatic test case generator tool that runs a benign version of an Android app and collects the set of calls made to sensitive APIs\kn{Is this the case for just Jamrozik's work or all of dynamic approaches. It is not clear}. This set of calls comprises a sandbox infrastructure. The sandbox might then monitor any call to sensitive APIs while the app is running, blocking calls that have not been identified during the exploratory phase---thereby protecting Android users from additional malicious behavior~\cite{DBLP:conf/icse/JamrozikSZ16}.
Jamrozik et al.~\cite{DBLP:conf/icse/JamrozikSZ16} argue in favor of dynamic analysis for mining sandboxes, instead of using static analysis---mostly because of the overapproximation problem: ``static analysis often assume that more behaviors are possible than actually would be''~\cite{DBLP:conf/icse/JamrozikSZ16}. In addition, code that uses dynamic features (such as reflection) poses additional challenges to static analysis algorithms---even though \emph{dynamic features} of programming languages are often used to introduce malicious behavior. Even though these claims are reasonable, previous research results do not present empirical assessments about the limitations of static analysis to mine sandboxes. Consequently, it is not clear whether and how both approaches (dynamic and static analysis) could complement each other in the process of mining Android sandboxes.

The same lack of understanding about the role of static analysis also appears in the work of Bao et al.~\cite{DBLP:conf/wcre/BaoLL18} (hereafter \blls), which presents an empirical study that compares different dynamic analysis tools for testing and mining Android sandboxes\kn{The use of word same here makes me wonder if readers could be confused. IMO the problem of how Static and dynamic analysis complement each other AND the role of static analysis in dynamic approaches are different (even if they sound similar), right? Reason I simply did not remove the word "same" is that I feel we need a bridge sentence here to make this transition of two similar sounding problems clear, because we only explore one of them.}. Their study leverages DroidFax~\cite{DBLP:conf/icsm/CaiR17a} to instrument $102$ pairs of Android apps (each pair comprising a benign and a malign version of an App) and to collect the information needed to mine sandboxes (that is, the calls to sensitive APIs).
Although the authors report a precision of at most 70\% of dynamic analysis tools to differentiate the benign and malicious versions of the apps, the authors ignore the fact that DroidFax statically analyzes the Android apps and also records calls to sensitive APIs (besides instrumenting the apps), which, as we discuss in this paper, leads to an overestimation of the performance of the dynamic analysis tools for mining sandboxes and might have introduced a possible threat to the conclusions of that work. Moreover, since DroidFax is an independent component (it is not part of the dynamic analysis tools explored in the \blls), a practitioner might run a dynamic analysis tool alone that is likely to present a smaller capacity to identify malware.

@@

\item A broad comprehension about the role of static analysis tools for mining
  sandboxes, showing that we can benefit from using both static and dynamic
  analysis for detecting malicious Android apps. In addition,
  we give evidence that a well known static analysis approach, based on
  taint analysis, leads to a performance similar to the dynamic analysis
  approach for diferenciating benign and malign versions of the same
  app (Section~\ref{sec:res-ss}).

>>>>

Almost two-thirds of the world use mobile technologies~\cite{Comscore}, and the Android Operating System has dominated the market of smartphones, tablets, and others electronic devices \cite{statcounter}. Due to this growing popularity, the number of incidents related to Android malicious software (malware) has significantly increased. Malware samples grow from just three families and 100 samples in 2010 to more than hundred families with thousands of samples in 2013~\cite{DBLP:journals/comsur/FarukiBLGGCR15,DBLP:journals/csur/SufatrioTCT15}%\kn{Can we here say by how much?}
. Security issues in Android software applications~\footnote{In this paper, we will use the terms Android Applications, Android Apps and Apps interchangeably, to represent Android software applications} have become a relevant research topic, and many techniques have been developed to identify vulnerabilities in Android apps~\cite{DBLP:conf/pldi/ArztRFBBKTOM14}, including the use of static analysis algorithms either to identify privacy leaks or to reveal the misuse of cryptographic primitives~\cite{krueger:ecoop-2018,rahaman:ccs-2019}, for instance.

Another perspective to the solving the same problem are approaches that rely on the use of dynamic analysis to mine Android sandboxes. Jamrozik et al. proposed BOXMATE tool~\cite{DBLP:conf/icse/JamrozikSZ16}, which with dynamic approaches, during an exploratory phase, one leverages an automatic test case generator tool that runs a benign version of an Android app and collects the set of calls made to sensitive APIs
%\kn{Is this the case for just Jamrozik's work or all of dynamic approaches. It is not clear}
. This set of calls comprises a sandbox infrastructure. The sandbox might then monitor any call to sensitive APIs while the app is running, blocking calls that have not been identified during the exploratory phase---thereby protecting Android users from additional malicious behavior~\cite{DBLP:conf/icse/JamrozikSZ16}.
Jamrozik et al. argue in favor of dynamic analysis for mining sandboxes, instead of using static analysis---mostly because of the overapproximation problem: ``static analysis often assume that more behaviors are possible than actually would be''~\cite{DBLP:conf/icse/JamrozikSZ16}. In addition, code that uses dynamic features (such as reflection) poses additional challenges to static analysis algorithms---even though \emph{dynamic features} of programming languages are often used to introduce malicious behavior. Even though these claims are reasonable, previous research results do not present empirical assessments about the limitations of static analysis to mine sandboxes. Consequently, it is not clear whether and how both approaches (dynamic and static analysis) could complement each other in the process of mining Android sandboxes.

The same lack of understanding about static and dynamic analysis complementing each other also appears in the work of Bao et al.~\cite{DBLP:conf/wcre/BaoLL18} (hereafter \blls), which presents an empirical study that compares different dynamic analysis tools for testing and mining Android sandboxes%\kn{The use of word same here makes me wonder if readers could be confused. IMO the problem of how Static and dynamic analysis complement each other AND the role of static analysis in dynamic approaches are different (even if they sound similar), right? Reason I simply did not remove the word "same" is that I feel we need a bridge sentence here to make this transition of two similar sounding problems clear, because we only explore one of them.}
. Their study leverages DroidFax~\cite{DBLP:conf/icsm/CaiR17a} to instrument $102$ pairs of Android apps (each pair comprising a benign and a malicious version of an App) and to collect the information needed to mine sandboxes (that is, the calls to sensitive APIs).
Although the authors report a precision of at most 70\% of dynamic analysis tools to differentiate the benign and malicious versions of the apps, the authors ignore the fact that DroidFax statically analyzes the Android apps and also records calls to sensitive APIs (besides instrumenting the apps), which, as we discuss in this paper, leads to an overestimation of the performance of the dynamic analysis tools for mining sandboxes and might have introduced a possible threat to the conclusions of that work. Moreover, since DroidFax is an independent component (it is not part of the dynamic analysis tools explored in the \blls), a practitioner might run a dynamic analysis tool alone that is likely to present a smaller capacity to identify malware.

@@

\item A broad comprehension about the role of static analysis tools for mining
  sandboxes, showing that we can benefit from using both static and dynamic
  analysis for detecting malicious Android apps. In addition,
  we give evidence that a well known static analysis approach, based on
  taint analysis, leads to a performance similar to the dynamic analysis
  approach for diferenciating benign and malicious versions of the same
  app (Section~\ref{sec:res-ss}).
